Supporting Information

### 2017-04-06 code/rmarkdown/text.Rmd
update benchmark anlysis
massive overhaul - packrat to manage r packages - reformat code according to lintr - fix book references

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

```{r, include=FALSE}
## initialization
# load .rda file
checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
session::restore.session('data/final/results.rda')

## preliminary processing
for (i in seq_len(nlayers(cs1.bioclim.RAST)))
  cs1.bioclim.RAST@layers[[i]]@file@name <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])

@@

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is represented by a single selected planning unit (ie. a $j \in J$ where $X_j = 1$). The degree to which a demand point $i$ is represented by a planning unit $j$ is determined by the distance between them ($d_{fsij}$). Generally--unless small space targets are used, eg. 0.000001, so that the solution is effectively unconstrained by the space target--demand points are represented by their closest selected planning units. In poorer quality solutions, demand points will be represented by planning units that have larger distances between them. As a consequence, the sum of the weighted distances between all of the demand demand points and the planning units used to represent them will be larger, and so, the solution will capture less of the variation described by the demand points.

@@

To showcase the behavior of the unreliable formulation, we conducted a simulation study and two case studies. To understand how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2). We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality.

## SIMULATION STUDY
### Methods
We simulated a hypothetical study area with square planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ grid (Fig. 2). We then simulated three species to inhabit this study area. Firstly, we simulated a hyper-generalist species (hereafter referred to as the "uniformly distributed species"). It all planning units with equal probability (Fig. 2a; eqn 4a). Secondly, we simulated a species with simple habitat requirements (hereafter referred to as the "normally distributed species"; Fig. 2b). This species was most likely to be found in planning units in the centre of the study area, and was simulated using density function of a multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b). Thirdly, we simulated a species two distinct populations (hereafter referred to as the "bimodally distributed species"; Fig. 2c). It was simulated using the maximum of two density of two multivariate normal distributions (eqn 4c). Planning unit occupancy was calculated using the $(X, Y)$ coordinates of the units' centroids and the corresponding equations. We used a geographic attribute space to provide an intuitive visualization of the solutions. Demand points were set as the planning units' centroids and were weighted according to the units' probability of occupancy.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.5 \tag*{eqn 4a} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{\matchcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
\matchcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\matchcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 4c} \\
%
\end{align*}

We generated four solutions for each species. Firstly, to represent solutions generated using conventional methods (eg. \texttt{Marxan}), we generated solutions using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets. Secondly, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % space targets to capture the geographic spread of each species (hereafter referred to as geographic spread targets). Thirdly, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`. Fourthly, to illustrate the combined effects of using amount and space targets as well as penalties for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic spread targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.

@@

The solution generated for the uniformly distributed species using an amount target prioritized `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area (Fig. 2a). For this species, all solutions containing this many planning units are optimal when using a `r sim.params.LST[[MODE]]$amount.target*100` % amount target to constrain the reserve selection problem. We obtained this particular solution due to artifacts in the solver. This solution captured a small amount of the geographic spread of the species (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled). In fact, the coverage was so poor that the proportion was negative because the solution captured a less representative sample than a solution containing one planning unit in the centre of the species' distribution.

By explicitly targeting representativeness, we obtained a solution that captured the geographic spread of the uniform species (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig. 3b). Although this solution secured an adequate amount of habitat and a representative sample of the species' geographic spread, this solution was highly fragmented. However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well-connected solution that met all of the objectives (Fig. 3d). This solution prioritized the same number of planning units as the previous solutions even though it is far superior. As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this objective.

@@

The solution generated for the normally distributed species using just an amount target prioritized planning units in the species' core range area (Fig. 3e). While this strategy is cost-effective for protecting habitat, it was ineffective for securing a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled) because it did not protect any planning units along the periphery of the species' distribution. By using amount and space targets, we were able to conserve an adequate amount of habitat and also secure a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig. 3f). Similar to the solutions for the uniformly distributed species, this solution was highly fragmented and we were able to obtain a better connected solution by specifying fragmentation penalties (Fig. 3h). However, unlike the solutions for the uniformly distributed species, the solution generated using an amount target, a space target, and fragmentation penalties required more planning units than the other solutions. These results suggest that solutions may need to prioritize more planning units to meet additional conservation objectives.

@@

The solution generated for the bimodally distributed species using an amount target only conserved individuals belonging to one of the two populations (Fig. 3i). As a consequence, this solution did not secure a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled). The addition of fragmentation penalties exacerbated this issue, and resulted in a solution that sampled even less of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig. 3k). However, this issue was resolved by using a space target  to obtain a solution that secured both populations (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % geographic spread secured; Fig. 3l). This finding suggests that species with large intra-specific variation could benefit the most from prioritizations generated using space-based targets.

## CASE STUDY 1
### Methods
We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the species' realized niches. By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience to environmental change [@r63]. We selected Queensland, Australia as the study area. We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1). We used the scores from the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation; Fig. 4).

We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_). These species span a range of different habitat requirements. We mapped the extent of occurrence for each species (Fig. 5). To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455]. We used this method because it is entirely reproducible using freely available data.

We generated `r cs1.params.LST[[MODE]]$dp.number` demand points for each species (Fig. S2). To ensure that the demand points reflected the core parts of the species' realized niches, we used the following method. For each species, we generated random points inside the species' geographic range and at each point extracted the principal component values at that location. We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches [implemented in the \texttt{hypervolume R} package; @r231]. We then generated uniformly distributed points inside the species' distribution in environmental space, and extracted the kernel density at their locations. These uniformly distributed points and associated density estimations were used as demand point coordinates and weights (respectively).

We generated two solutions. The first solution was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount targets for all species. The second solution was generated the same amount target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based target for each species.

### Results
Generally, the solution generated using amount targets preserved a representative sample of each the `r as.character(as.english(length(cs1.params.LST[[MODE]]$species.names)))` bird species' niches. (Figure S1). This solution secured over `r cs1.params.LST[[MODE]]$space.target*100` % of the realized niche of `r parsed.representative.space.held.names`. Yet it failed to achieve this for the `r parsed.not.representative.space.held.names`. This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of all features.

## CASE STUDY 2
### Methods
Here we used space-based targets to generate a prioritization securing a representative sample of a species' intra-specific genetic variation. We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462]. Although this dataset contains multiple species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}) as a simple case study, because it has been found to exhibit significant genetic structure [@r462]. Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Fig. 7a). They visited every second grid cell, and if the species was detected in a cell, samples were collected from three individuals. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci. In total, `r structurer::n.samples(cs2.spp.StructureData.LST[[1]])` individuals were genotyped at `r structurer::n.loci(cs2.spp.StructureData.LST[[1]])` markers.

We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented using the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458]. These continuous variables described the main axes of genetic variation within the species. We averaged together values corresponding to samples collected in the same grid cell, and used them to create a genetic attribute space (Figs. 7c and 7d). To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using scaled inverse great circle distances between the grid cells' centroids as weights [using the \texttt{ape R} package; @r464].

The grid cells were used as planning units for generating prioritizations. As previously mentioned, data were not collected in every grid cell, and so the planning units are arranged in a checkerboard pattern. The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit. Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points. To ensure that the solutions did not prioritize particularly costly areas, we obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell. We used these values to represent opportunity cost (Fig. 7b).

We generated two solutions. The first solution was generated using an `r cs2.params.LST[[MODE]]$amount.target*100` % amount-based target. The second solution was generated using the same amount-based target and an additional `r cs2.params.LST[[MODE]]$genetic.target*100` % genetic-based target.

@@

A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2.spp.nmds.LST[[1]]$stress,2)`; Figs. 7c--7d). Planning units located near each other contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$). In terms of the average genetic characteristics of individuals in the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c). This analysis supports previous work by Alvarez _et al._ [-@r461] who also found evidence of genetic structure within this species.

The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[1],2)` % sampled; Figs. 8a and 8c). This solution only conserved individuals in one of the two main genetic groups. In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fullfil the target. Whereas, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs. 8b and 8d). Although the solutions only differ by a single planning unit, by securing individuals in both main genetic groups the second solution was able to secure a much more representative sample of the species' genetic variation for only a minor increase in cost (`r round(cs2.spp.DF$Score[2],2)` compared to `r round(cs2.spp.DF$Score[1],2)` total cost).

>>>>

```{r, include = FALSE}
# set wd
knitr::opts_knit$set(root.dir = normalizePath("../.."))
```

```{r, include = FALSE}
## initialization
# load session
session::restore.session("data/final/results.rda")

## preliminary processing
for (i in seq_len(nlayers(cs1_bioclim_raster)))
  cs1_bioclim_raster@layers[[i]]@file@name <- paste0(getwd(),
    strsplit(cs1_bioclim_raster@layers[[i]]@file@name,
             "raptr-manuscript")[[1]][[2]])

@@

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is represented by a single selected planning unit (ie. a $j \in J$ where $X_j = 1$). The degree to which a demand point $i$ is represented by a planning unit $j$ is determined by the distance between them ($d_{fsij}$). Generally--unless near zero space targets are used so that the problem is effectively unconstrained by the target--demand points are represented by their closest selected planning units. In poorer quality solutions, demand points will be represented by planning units that have larger distances between them. As a consequence, the sum of the weighted distances between all of the demand demand points and the planning units used to represent them will be larger, and so, the solution will capture less of the variation described by the demand points.

@@

To showcase the behavior of the unreliable formulation, we conducted a simulation study and two case studies. To understand how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2). We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general_parameters$MIPGap * 100` % of optimality.

## SIMULATION STUDY
### Methods
We simulated a hypothetical study area with square planning units arranged in a $`r sqrt(sim_parameters$number_planning_units)` \times `r sqrt(sim_parameters$number_planning_units)`$ grid (Fig. 2). We then simulated three species to inhabit this study area. Firstly, we simulated a hyper-generalist species (hereafter referred to as the "uniformly distributed species"). It occupied all planning units with equal probability (Fig. 2a; eqn 4a). Secondly, we simulated a species with simple habitat requirements (hereafter referred to as the "normally distributed species"; Fig. 2b). This species was most likely to be found in planning units nearest to the center of the study area. It was simulated using the density function of a multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b). Thirdly, we simulated a species two distinct populations (hereafter referred to as the "bimodally distributed species"; Fig. 2c). It was simulated using the maximum density of two multivariate normal distributions (eqn 4c). For a given species, planning unit occupancy was calculated using the $(X, Y)$ coordinates of the units' centroids and the relevant equation. We used a geographic attribute space to provide an intuitive visualization of the solutions. Demand points were set as the planning units' centroids and were weighted according to the units' probability of occupancy.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.5 \tag*{eqn 4a} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
\mathcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 4c} \\
%
\end{align*}

We generated four solutions for each species. Firstly, to represent solutions generated using conventional methods (eg. \texttt{Marxan}), we generated solutions using `r sim_parameters$amount_target*100` % amount targets. Secondly, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and `r (sim_parameters$space_target*100)` % space targets to capture the geographic spread of each species (hereafter referred to as geographic spread targets). Thirdly, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and a boundary length modifier of `r sim_parameters$blm`. Fourthly, to illustrate the combined effects of using amount and space targets as well as penalties for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets, `r (sim_parameters$space_target*100)` % geographic spread targets, and boundary length modifiers of `r sim_parameters$blm`.

@@

The solution generated for the uniformly distributed species using an amount target prioritized `r dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$n` planning units near the southern end of study area (Fig. 2a). For this species, all solutions containing this number of planning units are optimal when using a `r sim_parameters$amount_target*100` % amount target. We obtained this particular solution due to artifacts in the solver (eg. seed for the random number generator, order of variables in the problem matrix). This solution captured a small amount of the geographic spread of the species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled). In fact, the coverage was so poor that the proportion was negative because the solution captured a less representative sample than a solution containing one planning unit in the center of the species' distribution.

By explicitly targeting representativeness, we obtained a solution that captured the geographic spread of the uniform species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig. 3b). Although this solution secured an adequate amount of habitat and a representative sample of the species' geographic spread, this solution was highly fragmented. However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well connected solution that met all of the objectives (Fig. 3d). This solution prioritized the same number of planning units as the previous solutions even though it is far superior. As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this property.

@@

The solution generated for the normally distributed species using just an amount target prioritized planning units in the species' core range area (Fig. 3e). While this strategy is cost-effective for protecting habitat, it was ineffective for securing a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled) because it did not protect any planning units along the periphery of the species' distribution. By using amount and space targets, we were able to conserve an adequate amount of habitat and also secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig. 3f). Similar to the solutions for the uniformly distributed species, this solution was highly fragmented and we were able to obtain a better connected solution by specifying fragmentation penalties (Fig. 3h). However, unlike the solutions for the uniformly distributed species, the solution generated using an amount target, a space target, and fragmentation penalties required more planning units than the other solutions. These results suggest that solutions may need to prioritize more planning units to meet additional conservation objectives.

@@

The solution generated for the bimodally distributed species using an amount target only conserved individuals belonging to one of the two populations (Fig. 3i). As a consequence, this solution did not secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled). The addition of fragmentation penalties exacerbated this issue, and resulted in a solution that sampled even less of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount target\nand BLM")$space_held, 2)` % sampled; Fig. 3k). However, this issue was resolved by using a space target  to obtain a solution that secured both populations (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % geographic spread secured; Fig. 3l). This finding suggests that species with large intra-specific variation could benefit the most from prioritizations generated using space-based targets.

## CASE STUDY 1
### Methods
We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the species' realized niches. By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience to environmental change [@r63]. We selected Queensland, Australia as the study area. We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1). We used the scores from the first `r as.character(english::as.english(cs1_parameters$number_pc_layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1_pca_data[[4]][cs1_parameters$number_pc_layers], 1)` % of the total variation; Fig. 4).

We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_). These species span a range of different habitat requirements. We mapped the extent of occurrence for each species (Fig. 5). To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455]. We used this method because it is entirely reproducible using freely available data.

We generated `r cs1_parameters$dp_number` demand points for each species (Fig. S2). To ensure that the demand points reflected the core parts of the species' realized niches, we used the following method. For each species, we generated random points inside the species' geographic range and at each point extracted the principal component values at that location. We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species using a manually chosen bandwidth of `r cs1_parameters$dp_bandwidth[1]` and a `r cs1_parameters$dp_quantile` quantile to map the core parts of the species' niches [implemented in the \texttt{hypervolume R} package; @r231]. We then generated uniformly distributed points inside the species' distribution in environmental space, and extracted the kernel density at their locations. These uniformly distributed points and associated density estimations were used as demand point coordinates and weights (respectively).

We generated two multi-species prioritizations. The first solution was generated using `r cs1_parameters$amount_target*100` % amount targets for each species. The second solution was generated the same amount targets with an additional `r cs1_parameters$space_target*100` % niche-based target for each species.

### Results
Generally, the solution generated using just amount targets preserved a representative sample of each the `r as.character(english::as.english(length(cs1_parameters$species_names)))` bird species' niches. (Figure S1). This solution secured over `r cs1_parameters$space_target*100` % of the realized niche of `r parsed_representative_space_held_names`. Yet it failed to achieve this for the `r parsed_not_representative_space_held_names`. This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of all features.

## CASE STUDY 2
### Methods
Here we used space-based targets to generate a prioritization securing a representative sample of a species' intra-specific genetic variation. We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462]. Although this dataset contains multiple species, we used data for the `r tolower(cs2_parameters$common_name)` (\textit{`r gsub('_', ' ',cs2_parameters$species_name,fixed=TRUE)`}) as a simple case study, because exhibits significant inter-population genetic structure [@r462]. Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Fig. 7a). They visited every second grid cell, and if the species was detected in a cell, samples were collected from three individuals. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci. In total, `r structurer::n.samples(cs2_spp_structure_data[[1]])` individuals were genotyped at `r structurer::n.loci(cs2_spp_structure_data[[1]])` markers.

We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented using the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458]. These continuous variables described the main axes of genetic variation within the species. We averaged together values corresponding to samples collected in the same grid cell, and used them to create a genetic attribute space (Figs. 7c and 7d). To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using scaled  great circle distances between the grid cells' centroids [using the \texttt{ape R} package; @r464].

The grid cells were used as planning units for generating prioritizations. Since data were not collected in every grid cell, and the planning units were arranged in a checkerboard pattern. The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit. Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points. To ensure that the solutions did not prioritize particularly costly areas, we obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell. We used these values to represent opportunity cost (Fig. 7b).

We generated two solutions. The first solution was generated using an `r cs2_parameters$amount_target*100` % amount-based target. The second solution was generated using the same amount-based target and an additional `r cs2_parameters$genetic_target*100` % genetic-based target.

@@

A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2_spp_nmds[[1]]$stress,2)`; Figs. 7c--7d). Planning units located near each other contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1_nmds_morans_i[[1]][[1]]$observed,2)`$, $P `r lazyWeave::pvalString(cs1_nmds_morans_i[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1_nmds_morans_i[[1]][[2]]$observed,2)`$, $P `r lazyWeave::pvalString(cs1_nmds_morans_i[[1]][[2]]$p.value)`$). In terms of the average genetic characteristics of individuals in the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c). This analysis supports previous work by Alvarez _et al._ [-@r461] who also found evidence of genetic structure within this species.

The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2_spp_results$genetic[1], 2)` % sampled; Figs. 8a and 8c). This solution only conserved individuals in one of the two main genetic groups. In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fullfil the target. Whereas, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2_spp_results$genetic[2],2)` % sampled; Figs. 8b and 8d). Although the solutions only differ by a single planning unit, by securing individuals in both main genetic groups the second solution was able to secure a much more representative sample of the species' genetic variation for only a minor increase in cost (`r round(cs2_spp_results$Score[2],2)` compared to `r round(cs2_spp_results$Score[1],2)` total cost).
